SpikeIn MSstatsPTM Analysis

### 2022-08-09 code/Spike-in_Vizualization.Rmd
implementation and eval updates

plot_df <- rbind(model_df$ADJUSTED.Model %>% select(-GlobalProtein), 
              model_df$PTM.Model %>% select(-c(MissingPercentage,ImputationPercentage, issue)))
plot_df$protadj <- factor(plot_df$protadj, levels = c('no adjustment','protein adjustment'))

@@

# Volcano plot
plot_df %>% arrange(labeling) %>% 
  filter(abs(log2FC) < Inf) %>% 
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,5)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "Proposed Method VolcanoPlot")
```

@@

```{r, echo = FALSE, fig.width = 10, fig.height = 5}
unadj_spike <- model_df$PTM.Model %>% filter(labeling == "heavy")
unadj_spike$Adjustment <- "no adjustment"
adj_spike <- model_df$ADJUSTED.Model %>% filter(labeling == "heavy")
adj_spike$Adjustment <- "protein adjustment"
spike <- rbindlist(list(unadj_spike, adj_spike), fill = TRUE)

@@

plot_df <- rbind(limma_model$ADJUSTED.Model, limma_model$PTM.Model, fill = TRUE)
plot_df$protadj <- factor(plot_df$protadj, levels = c('no adjustment','protein adjustment'))

@@

# Volcano plot
plot_df %>% arrange(labeling) %>% 
  filter(is.finite(Log2FC)) %>%
  ggplot(aes(Log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  # scale_y_continuous(limits = c(0,3)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "Limma VolcanoPlot")
```

@@

proposed_plot_df$Model <- "Proposed Method"
limma_plot_df$Model <- "Limma"
limma_plot_df$Protein <- limma_plot_df$PTM
limma_plot_df$log2FC <- limma_plot_df$Log2FC

combo_plot_df <- rbind(proposed_plot_df[, c("Protein", "Label", "log2FC", 
                                            "adj.pvalue", "labeling", 
                                            "trueLog2FC", "contrast_truth", 
                                            "protadj", "Model")], 
                       limma_plot_df[, c("Protein", "Label", "log2FC", 
                                            "adj.pvalue", "labeling", 
                                            "trueLog2FC", "contrast_truth", 
                                            "protadj", "Model")])
combo_plot_df$Model <- factor(combo_plot_df$Model, levels=c('Proposed Method','Limma'))

combo_plot_df %>% arrange(Model, labeling) %>% 
  filter(abs(log2FC) < Inf & Label %in% c("mix3 vs mix2",
                                          "mix4 vs mix2")) %>%
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ Model + contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,3)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "SpikeIn VolcanoPlot", x= "Log2FC")

@@

@@

# Volcano plot
png("../supplementary/sim_new/spike_in_anova_volcano.png", width = 1000, height = 500)
anova_results %>% arrange(labeling) %>% 
  filter(is.finite(log2FC)) %>%
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.1, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 14), 
       axis.text.y = element_text(size = 14), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  # scale_y_continuous(limits = c(0,1.5)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) + 
  labs(title = "ANOVA VolcanoPlot")
dev.off()

@@

spike <- rbindlist(list(unadj_spike_msstats, adj_spike_msstats, unadj_spike_ttest,
                        adj_spike_ttest, unadj_spike_limma, adj_spike_limma), fill = TRUE)
spike <- as.data.table(spike)
spike[spike$Model == "MSstatsPTM", Model := "Proposed"]
#plot_df[plot_df$Label == 'mix1-mix3'][['contrast_truth']]
spike$`color_label` <- "True LogFC"

png("../main/images/spike_in_fc.png", width = 750, height = 500)
spike %>% mutate(Comparison = str_split(contrast_truth, " ", simplify = TRUE)[,1],
      Model = factor(Model, levels=c("Proposed", "Limma", "ANOVA"))) %>%
  filter(!Label %in% c("mix1-mix2", "mix3-mix4")) %>% 
  ggplot() + 
  geom_boxplot(aes(x = Label, y = log2FC, fill = Model)) + 
  geom_point(aes(y = trueLog2FC, x = Label, color = color_label), 
             shape = 4, size = 4, stroke = 2) + 
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(name = "", values="red", label="True LogFC") + 
 facet_wrap(.~Adjustment) + 
    theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust = 1, size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=18),
       axis.title.y = element_text(size = 18),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 18),
       legend.position = "bottom") + 
  labs(title = "Fold Change of Spike-in Peptides", x = "Comparison with Expected FC")
dev.off()

>>>>

model_df$ADJUSTED.Model$protadj = "MSstatsPTM"
model_df$PTM.Model$protadj = "MSstats"

plot_df <- rbind(model_df$ADJUSTED.Model %>% select(-GlobalProtein), 
              model_df$PTM.Model %>% select(-c(MissingPercentage,ImputationPercentage, issue)))
plot_df$protadj <- factor(plot_df$protadj, levels = c('MSstats','MSstatsPTM'))

@@

# Volcano plot
png("../supplementary/sim_new/spike_in_msstatsptm_volcano.png", width = 1000, height = 500)
plot_df %>% arrange(labeling) %>% 
  filter(abs(log2FC) < Inf) %>% 
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,5)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "MSstatsPTM VolcanoPlot")
dev.off()
```

@@

```{r, echo = FALSE, fig.width = 10, fig.height = 5}
unadj_spike <- model_df$PTM.Model %>% filter(labeling == "heavy")
unadj_spike$Adjustment <- "MSstats"
adj_spike <- model_df$ADJUSTED.Model %>% filter(labeling == "heavy")
adj_spike$Adjustment <- "MSstatsPTM"
spike <- rbindlist(list(unadj_spike, adj_spike), fill = TRUE)

@@

limma_model$ADJUSTED.Model$protadj = "Limma"
limma_model$PTM.Model$protadj = "basic Limma"

plot_df <- rbind(limma_model$ADJUSTED.Model, limma_model$PTM.Model, fill = TRUE)
plot_df$protadj <- factor(plot_df$protadj, levels = c('basic Limma','Limma'))

@@

# Volcano plot
png("../supplementary/sim_new/spike_in_limma_volcano.png", width = 1000, height = 500)
plot_df %>% arrange(labeling) %>% 
  filter(is.finite(Log2FC)) %>%
  ggplot(aes(Log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  # scale_y_continuous(limits = c(0,3)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "Limma VolcanoPlot")
dev.off()
```

@@

proposed_plot_df$Model <- "MSstatsPTM"
proposed_plot_df[proposed_plot_df$protadj == "MSstats"]$protadj = "no adjustment"
proposed_plot_df[proposed_plot_df$protadj == "MSstatsPTM"]$protadj = "protein adjustment"
limma_plot_df$Model <- "Limma"
limma_plot_df$Protein <- limma_plot_df$PTM
limma_plot_df$log2FC <- limma_plot_df$Log2FC
limma_plot_df[limma_plot_df$protadj == "basic Limma"]$protadj = "no adjustment"
limma_plot_df[limma_plot_df$protadj == "Limma"]$protadj = "protein adjustment"

combo_plot_df <- rbind(proposed_plot_df[, c("Protein", "Label", "log2FC", 
                                            "adj.pvalue", "labeling", 
                                            "trueLog2FC", "contrast_truth", 
                                            "protadj", "Model")], 
                       limma_plot_df[, c("Protein", "Label", "log2FC", 
                                            "adj.pvalue", "labeling", 
                                            "trueLog2FC", "contrast_truth", 
                                            "protadj", "Model")])
combo_plot_df$Model <- factor(combo_plot_df$Model, levels=c('MSstatsPTM','Limma'))

png("../main/images/spike_in_volcano.png", width = 750, height = 500)
combo_plot_df %>% arrange(Model, labeling) %>% 
  filter(abs(log2FC) < Inf & Label %in% c("mix3 vs mix2",
                                          "mix4 vs mix2")) %>%
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.2, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ Model + contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  scale_y_continuous(limits = c(0,3)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(title = "SpikeIn VolcanoPlot", x= "Log2FC")
dev.off()

@@

anova_results[anova_results$protadj == "no adjustment"]$protadj = "basic ANOVA"
anova_results[anova_results$protadj == "protein adjustment"]$protadj = "ANOVA"

@@

# Volcano plot
png("../supplementary/sim_new/spike_in_anova_volcano.png", width = 1000, height = 500)
anova_results %>% arrange(protadj, labeling) %>% 
  filter(is.finite(log2FC)) %>%
  ggplot(aes(log2FC, -log10(adj.pvalue))) + 
  geom_point(aes(color = labeling, alpha = labeling), size = 2.2) + 
  scale_color_manual(values = c("#999999", "#D55E00")) + 
  scale_alpha_manual(values = c(0.1, 1)) + 
  #scale_size_manual(values = c(2, 1)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_vline(xintercept = 0) + 
  geom_vline(aes(xintercept = trueLog2FC),color = "darkred", linetype= "dashed") +
  geom_hline(yintercept = -log10(0.05), color = "darkred") + 
  facet_grid(protadj ~ contrast_truth) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 14), 
       axis.text.y = element_text(size = 14), 
       legend.text=element_text(size=16),
       axis.title.y = element_text(size = 18, vjust=2),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 15),
       legend.title =  element_blank(),
       legend.position = "bottom",
       legend.box.margin=margin(-10,-10,-10,-10)) + 
  # scale_y_continuous(limits = c(0,1.5)) + 
  guides(colour = guide_legend(override.aes = list(size=5))) + 
  labs(title = "ANOVA VolcanoPlot")
dev.off()

@@

spike <- rbindlist(list(unadj_spike_msstats, adj_spike_msstats, unadj_spike_ttest,
                        adj_spike_ttest, unadj_spike_limma, adj_spike_limma), fill = TRUE)
spike <- as.data.table(spike)
# spike[spike$Model == "MSstatsPTM", Model := "Proposed"]
#plot_df[plot_df$Label == 'mix1-mix3'][['contrast_truth']]
spike$`color_label` <- "True LogFC"

png("../main/images/spike_in_fc.png", width = 750, height = 500)
spike %>% mutate(Comparison = str_split(contrast_truth, " ", simplify = TRUE)[,1],
      Model = factor(Model, levels=c("MSstatsPTM", "Limma", "ANOVA"))) %>%
  filter(!Label %in% c("mix1-mix2", "mix3-mix4")) %>% 
  ggplot() + 
  geom_boxplot(aes(x = Label, y = log2FC, fill = Model)) + 
  geom_point(aes(y = trueLog2FC, x = Label, color = color_label), 
             shape = 4, size = 4, stroke = 2) + 
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(name = "", values="red", label="True LogFC") + 
 facet_wrap(.~Adjustment) + 
    theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust = 1, size = 16), 
       axis.text.y = element_text(size = 16), 
       legend.text=element_text(size=18),
       axis.title.y = element_text(size = 18),
       axis.title.x = element_text(size = 18),
       title = element_text(size = 20),
       strip.text = element_text(size = 18),
       legend.position = "bottom") + 
  labs(title = "Fold Change of Spike-in Peptides", x = "Comparison with Expected FC")
dev.off()
